---
title: "hw3"
author: "Hao Zheng(hz2772)"
output: github_document
---
library packages and set plot size
```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(p8105.datasets)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1 

**Q:** Description of "instacart" data
```{r, message = FALSE, results='hide'}
data("instacart")
## description
skimr::skim(instacart)
```
**A:** This dataset has 1384617 rows and 15 columns, and 4 variables are character type, like eval_set and product_name, and 11 variables are numeric type, like order_id and product_id. We can observe from the data that the product whose id is 49683 has been purchased `r instacart %>% filter(product_id==49683) %>% count()` times.


**Q:** How many aisles are there, and which aisles are the most items ordered from?
```{r}
instacart %>% 
  count(aisle,sort = TRUE) %>%
  base::nrow() # return the number of rows
instacart %>% 
  count(aisle,sort = TRUE) %>%
  slice(1:1) %>%
  select(aisle)
```
**A:** There are `r instacart %>% count(aisle,sort = TRUE) %>%base::nrow()` aisles, and `r instacart %>% count(aisle,sort = TRUE) %>% slice(1:1) %>% select(aisle)` are the most items ordered from.


**Q:** Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
instacart %>%
  count(aisle,sort = TRUE) %>%
  filter(n >= 10000) %>%
  rename(number = n) %>%
  mutate(aisle = fct_relevel(aisle, c(aisle))) %>%
  ggplot(aes(x = aisle, y = number, fill = rainbow(39))) + 
  geom_bar(stat = "identity") + 
  guides(fill=FALSE) +
  coord_flip() +
  labs(
    title = "Number of items ordered in each aisle",
    x = "Aisle",
    y = "Number of items ordered",
    caption = "Data from the p8105.datasets package"
  ) +
  theme_bw() 
```


**A:** The outcome is the plot.


**Q:** Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
baking_count = 
  instacart %>%
  filter(aisle == "baking ingredients") %>%
  count(product_name,aisle,sort = TRUE) %>%
  slice(1:3)

dogfd_count = 
  instacart %>%
  filter(aisle == "dog food care") %>%
  count(product_name,aisle,sort = TRUE) %>%
  slice(1:3)
  
pcdvg_count = 
  instacart %>%
  filter(aisle == "packaged vegetables fruits") %>%
  count(product_name,aisle,sort = TRUE) %>%
  slice(1:3)

three_aisles = 
  rbind(
    baking_count,
    dogfd_count,
    pcdvg_count
  ) %>%
  rename(number = n) 

three_aisles %>%
  knitr::kable()
```


**A:** The outcome is the table.


**Q:** Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) %>%
  select(product_name,order_hour_of_day,order_dow) %>%
  group_by(product_name,order_dow) %>%
  summarise(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_hour"
  ) %>%
  knitr::kable()
```


**A:** The outcome is the table.


## Problem 2

**Q:** Data cleaning
```{r}
data("brfss_smart2010") 
brfss = 
  brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  mutate(response = forcats::fct_relevel(response, c("Poor","Fair","Good","Very good","Excellent")))
```


**Q:** In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r, message=FALSE}
brfss %>%
  select(year, locationabbr, locationdesc) %>%
  filter(year == 2002) %>%
  group_by(year,locationabbr) %>%
  summarise(state_location = n_distinct(locationdesc)) %>%
  filter(state_location >= 7) %>%
  select(locationabbr) %>%
  c()
brfss %>%
  select(year, locationabbr, locationdesc) %>%
  filter(year == 2010) %>%
  group_by(year,locationabbr) %>%
  summarise(state_location = n_distinct(locationdesc)) %>%
  filter(state_location >= 7) %>%
  select(locationabbr) %>%
  c()
```
**A:** In 2002, states that were observed at 7 or more locations are "CT", "FL", "MA", "NC", "NJ" and "PA". In 2010, states that were observed at 7 or more locations are "CA", "CO", "FL", "MA", "MD", "NC" ,"NE" ,"NJ", "NY", "OH", "PA", "SC", "TX" and "WA". 


**Q:** Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}
brfss %>%
  filter(response == "Excellent") %>%
  drop_na(data_value) %>%
  select(year, locationabbr, locationdesc, response, data_value) %>%
  group_by(year, locationabbr) %>%
  summarise(state_mean = mean(data_value)) %>%
  ## plot
  ggplot(aes(x = year, y = state_mean, color = locationabbr))+
  geom_line(alpha = .522, size =0.522) + 
  labs(
    title = "Mean data value plot",
    x = "Year",
    y = "Mean data value"
  ) + 
  theme_classic() + 
  theme(legend.position = "right", legend.key.size=unit(2,'mm'))

```


**A:** The plot shows lines of mean data_value for each state across years.


**Q:** Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss %>%
  filter(response %in% c("Poor","Fair","Good","Very good","Excellent"), 
         year %in% c(2006,2010),
         locationabbr == "NY"
  ) %>%
  group_by(response) %>%
  ggplot(aes(x = data_value, fill = response)) + 
  geom_density(alpha = .4, adjust = .5, color = "pink") + 
  scale_x_continuous(limits = c(0,45)) +
  labs(
    title = "Data value distribution plot",
    x = "Data value",
    y = "Density"
  ) +  
  facet_grid(. ~ year) +
  theme_classic() + 
  theme(legend.position = "right")
```


**A:** The plot shows the distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.


## Problem 3

**Q:** Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
```{r, message = FALSE, results='hide'}
accel_df = 
  read.csv("data/accel_data.csv") %>%
  janitor::clean_names() %>%
  mutate(weekday_vs_weekend = ifelse(day %in% c("Saturday","Sunday"), "weekend", "weekday")) %>%   
  mutate(day = fct_relevel(day, c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))) 
## description
skimr::skim(accel_df)
``` 
**A:** This dataset has 35 rows and 1444 columns, and weekday_vs_weekend variables is character type, day variables is character type and 1442 variables are numeric type, like week, day_id and activity_*. We can observe from the data that the activity_2 value of the week 1 and day_id==5 is `r accel_df %>% filter(week==1, day_id==5) %>% select(activity_2)`.


**Q:** Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. 
```{r}
accel_df %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "value"
  ) %>%
  mutate(minute = as.integer(minute)) %>%
  group_by(week, day_id) %>%
  summarise(total_act = sum(value))%>%
  knitr::kable()
```
**A:** The activity usually will rise at the final day of a week.


**Q:** Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.
```{r}
accel_df %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "value"
  ) %>%
  mutate(minute = as.integer(minute)) %>%
  group_by(day_id, day) %>% 
  ggplot(
    aes(x = minute, y = value, color = day)
  ) +
  geom_line(alpha = .522) +
  scale_x_discrete(
    limits = c(0:1440),
    breaks = c(1,181,361,541,721,901,1081,1261,1440), 
    labels = c("00","03","06","09","12","15","18","21","24")) + 
  labs(
    title = "Activity plot of each day",
    x = "Hour",
    y = "Activity"
  ) 
```


**A:** The graph shows the activity is not active during 00~06 AM, and is active during 09-12 O'clock and 19-22 O'clock. The activity is active in Tuesday and Sunday, and the activity of Saturday is almost maintaining a fixed level. 






















































