---
title: "hw3"
author: "Hao Zheng(hz2772)"
output: github_document
---
library packages and set plot size
```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(p8105.datasets)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1 

**Q:** Description of "instacart" data
```{r, message = FALSE, results='hide'}
data("instacart")
## description
skimr::skim(instacart)
```
**A:** This dataset has 1384617 rows and 15 columns, and 4 variables are character type, like eval_set and product_name, and 11 variables are numeric type, like order_id and product_id. We can observe from the data that the product whose id is 49683 has been purchased `r instacart %>% filter(product_id==49683) %>% count()` times.


**Q:** How many aisles are there, and which aisles are the most items ordered from?
```{r}
instacart %>% 
  count(aisle,sort = TRUE) %>%
  base::nrow() # return the number of rows
instacart %>% 
  count(aisle,sort = TRUE) %>%
  slice(1:1) %>%
  select(aisle)
```
**A:** There are `r instacart %>% count(aisle,sort = TRUE) %>%base::nrow()` aisles, and `r instacart %>% count(aisle,sort = TRUE) %>% slice(1:1) %>% select(aisle)` are the most items ordered from.


**Q:** Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
instacart %>%
  count(aisle,sort = TRUE) %>%
  filter(n >= 10000) %>%
  rename(number = n) %>%
  mutate(aisle = fct_relevel(aisle, c(aisle))) %>%
  ggplot(aes(x = aisle, y = number, fill = rainbow(39))) + 
  geom_bar(stat = "identity") + 
  guides(fill=FALSE) +
  coord_flip() +
  labs(
    title = "Number of items ordered in each aisle",
    x = "Aisle",
    y = "Number of items ordered",
    caption = "Data from the p8105.datasets package"
  ) +
  theme_bw() 
```


**A:** The outcome is the plot.


**Q:** Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
baking_count = 
  instacart %>%
  filter(aisle == "baking ingredients") %>%
  count(product_name,aisle,sort = TRUE) %>%
  slice(1:3)

dogfd_count = 
  instacart %>%
  filter(aisle == "dog food care") %>%
  count(product_name,aisle,sort = TRUE) %>%
  slice(1:3)
  
pcdvg_count = 
  instacart %>%
  filter(aisle == "packaged vegetables fruits") %>%
  count(product_name,aisle,sort = TRUE) %>%
  slice(1:3)

three_aisles = 
  rbind(
    baking_count,
    dogfd_count,
    pcdvg_count
  ) %>%
  rename(number = n) 

three_aisles %>%
  knitr::kable()
```


**A:** The outcome is the table.


**Q:** Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) %>%
  select(product_name,order_hour_of_day,order_dow) %>%
  group_by(product_name,order_dow) %>%
  summarise(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_hour"
  ) %>%
  knitr::kable()
```


**A:** The outcome is the table.
























































